### 2023-09-26 manuscript.tex
Merge pull request #6 from MMV-Lab/resubmission [ci skip]
This build is based on
https://github.com/MMV-Lab/im2im-paper/commit/95b4e4177071248e2ce1c3ffec17cc0c9df52191.

This commit was created by the following CI build and job:
https://github.com/MMV-Lab/im2im-paper/commit/95b4e4177071248e2ce1c3ffec17cc0c9df52191/checks
https://github.com/MMV-Lab/im2im-paper/actions/runs/6311060347

This manuscript
(\href{https://MMV-Lab.github.io/im2im-paper/v/e52fbde2994a307c76a8d2ec15346bfec89f46a9/}{permalink})
was automatically generated
from \href{https://github.com/MMV-Lab/im2im-paper/tree/e52fbde2994a307c76a8d2ec15346bfec89f46a9}{MMV-Lab/im2im-paper@e52fbde}
on June 19, 2023.

@@

\emph{Bottleneck: existing codd not easy to understand or to extend or re-use}.

@@

Moreover, as ML scientists, have you ever overwhelmed by different training tricks for different methods or been curious about if certain state-of-the-art training methods can boost the performance of existing models? With the pytorch-lightning backend, \emph{MMV\_Im2Im} allows you to enjoy different state-of-the-art ML engineering techniques without changing the code, e.g., stochastic weight averaging {[}\protect\hyperlink{ref-qzeQFRn9}{15}{]}, single precision training, automatic batch size determination, different optimizers, different learning rate schedulers, easy deployment on different devices, distributed training on multi-GPU (even multi-node), logging with common loggers such as Tensorboard, etc.. In short, with the pytorch-lightning based universal boilerplate, bioimaging researchers can really focus on research and develop novel methods for their biomedical applications, without worrying about the ML engineering works (which are usually lack in non-computer-science labs).

@@

The toolbox is designed for both people with or without extensive experience with ML and Python programming. It is not rare to find biomedical image analysis software that are very easy to use on a set of problems, but very hard to extend or adjust to other different but essentally related problems, or find some with great flexibility with tubable knobs at all levels, but unfortunately not easy for unexperienced users. To address this issue, we design the toolbox in a systematically modularized way with various levels of configurability. One can use the toolbox with a single command as simple as \texttt{run\_im2im\ -\/-config\ train\_semanticseg\_3d\ -\/-data.data\_path\ /path/to/data} or make customization on details directly from a human-readable configuration file, such as choosing batch normalization or instance normalization in certain layers of the model, or adding extra data augmentation steps, etc.. We provide an extensive list of more than 20 example configurations for various appplications and comprehensive documentation to address common questions for users as reference. For users preferring graphical interface, another napari plugin for MMV toolbox has been planned as the extension of \emph{MMV\_Im2Im} (see Discussion for details).

@@

Finally, we showed the comparison between three different types of models, an FCN-type model (i.e., fnet), a pix2pix-type model, and a cycleGAN-type model. For fair comparison, we used fnet as the same backbone for all three types of models. In theory, the pix2pix-type model can be trained in two different ways: from scratch or initializing the generator with a pre-trained fnet (trained as FCN). Examples of the comparison results were shown in the last two rows in Figure \ref{fig:labelfree}-B. Visually, it is evident that the additional adversarial components (i.e., the discriminator) could generate images with more realistic apprearnce than a typical FCN-type model alone, but again, we leave the appropriate quantitative evaluations to users' specfic biomedical studies.

@@

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2500}}@{}}
\caption{Evaluation of the final 3D label-free models for four different nuclear structures. \label{tbl:labelfree_table}}\label{tbl:labelfree_table}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Dataset
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Pearson Correlation
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Structural Similarity
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\# of Test Data
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Dataset
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Pearson Correlation
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Structural Similarity
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\# of Test Data
\end{minipage} \\
\midrule()
\endhead
FBL & 0.864 ± 0.021 & 0.831 ± 0.034 & 50 \\
HIST1H2BJ & 0.825 ± 0.034 & 0.675 ± 0.073 & 55 \\
LMNB1 & 0.853 ± 0.027 & 0.669 ± 0.059 & 50 \\
NPM1 & 0.912 ± 0.015 & 0.795 ± 0.039 & 55 \\
\bottomrule()
\end{longtable}

\begin{figure}
\hypertarget{fig:labelfree}{%
\centering
\includegraphics[width=0.76\textwidth,height=0.9\textheight]{images/labelfree.png}
\caption{Examples of labelfree results. p/c/s refers percentile normalization, center normalization, and standard normalization, respectively (see main text for details).}\label{fig:labelfree}
}
\end{figure}

\hypertarget{d-semantic-segmentation-of-tissues-from-he-images}{%
\subsubsection{2D semantic segmentation of tissues from H\&E images}\label{d-semantic-segmentation-of-tissues-from-he-images}}

Segmentation is a common image processing task, and can be considered as a special type of image-to-image transformation, where the generated images are segmentation masks. DL based methods have achieved huge success in 2D semantic segmentation in biomedical images. In this example, we demonstrated \emph{MMV\_Im2Im} on a pathology application to segment glands from hematoxylin and eosin (H\&E) stained tissue images from the 2015 Gland Segmentation challenge {[}\protect\hyperlink{ref-45Sirz1X}{29},\protect\hyperlink{ref-XAffSYIR}{30}{]}. Stain normalization is an important pre-processing step in order to develop models robust to stain variation and tissue variations. \emph{MMV\_Im2Im} included a classic stain normalization method {[}\protect\hyperlink{ref-tQhnZyjK}{31}{]} as a pre-processing step. The effect of stain normalization can be observed in Figure \ref{fig:2d_gland}-A and B. We trained a simple attention UNet model {[}\protect\hyperlink{ref-OCow1hly}{26}{]}. Evaluated on the two different hold-out test sets, the model achieved F1-score, 0.90 and 0.86 on test set A and test set B, respectively. The performance was competitive comparing to the methods reported in the challenge report {[}\protect\hyperlink{ref-XAffSYIR}{30}{]}, especially with much more consistent performance across the two different test sets. Example results can be found in Figure \ref{fig:2d_gland}-C.

\begin{figure}
\hypertarget{fig:2d_gland}{%
\centering
\includegraphics[width=0.75\textwidth,height=\textheight]{images/gland.png}
\caption{Example results of 2D semantic segmentation of gland in H\&E images.}\label{fig:2d_gland}
}
\end{figure}

@@

Instance segmentation is a type of segmentation problem that goes beyond semantic segmentation. The goal is to differentiate not only between different types of objects, but also different instances of the same type of objects. Currently, the \emph{MMV\_Im2Im} package supports \emph{EmbedSeg}-type models. The major benefit of EmbedSeg-type models is their agnosticism to the morphology and dimensionality of the object instances, compared to other models such as StarDist {[}\protect\hyperlink{ref-tIIG2f8K}{32},\protect\hyperlink{ref-14h90Vfg0}{33}{]}, SplineDist {[}\protect\hyperlink{ref-17Yrl6WGQ}{34}{]} and CellPose {[}\protect\hyperlink{ref-TugPkOLy}{35}{]}. For example, different from the others, \emph{EmbedSeg}-type models are even able to generate instance segmentation where each instance contains multiple connected components. Additional frameworks such as OmniPose {[}\protect\hyperlink{ref-lXzmjM5n}{36}{]} will be supported in future versions. Another mainstream category of instance segmentation methods are detection-based models, such as Mask-RCNN {[}\protect\hyperlink{ref-xi8wnibR}{37}{]}. However, these models are better suited to the detection framework rather than image-to-image transformation (see Discussion section for details).

The \emph{EmbedSeg}-type models were re-implemented according to the original paper {[}\protect\hyperlink{ref-K2ugNcVa}{10},\protect\hyperlink{ref-QmYuUQ5K}{11}{]} following the generic boilerplate in \emph{MMV\_Im2Im}, with significant improvment. First of all, following the modular design of \emph{MMV\_Im2Im}, it is flexible to use different neural network models as the backbone. For 3D anisotropic microscopy image, the original backbone ERFNet {[}\protect\hyperlink{ref-XAkgs3Nh}{38}{]} doesn't take the anisotropic dimensions into account and therefore may not perform well or even not applicable. In this scenario, it is straighforward to employ another anisotropic neural network bone, such as the anisotropic U-Net in {[}\protect\hyperlink{ref-jM3v1UjQ}{19}{]} or the anisotropic version of Dynamic U-Net in MONAI. Second, we significantly improve training strategy. The original version requires pre-cropping patches centered around each instance and pre-calculated the center images and class images. This may generate a massive amount of additional data on the disk. More importantly, such pre-cropping makes data augmentation nearly impossible, except the simple ones like flipping (otherwise, the pre-calculated centers might be wrong), and also greatly undersamples around negative cases (e.g., background). For example, we have observed that for an Embedseg model training only with patches centered around instances, the model may suffer from degraded performance during inference when there are a large amound of background areas without any instances. Again, following the modular design of \emph{MMV\_Im2Im}, it is now possible to do on-the-fly data augmentation and patch sampling, even weighted patch sampling. Third, our improved \emph{Embedseg}-type models can accept an exclusion mask so that certain parts of the images can be ignored during training. This is especially useful for partially annotated ground truth. For large images, it could be extremely time-consuming to require every single instance to be annotated. The exclusion masks can address this bottleneck. Another extension comparing to the original implementation was that the \emph{MMV\_Im2Im} package made sliding windowing inference straightforward, and therefore permitted easy handling of images of any size during inference.

@@

For the 2D problem, we adopted the same network backbone as in the original \emph{EmbedSeg} paper. Example results on a small holdout set are shown in Figure \ref{fig:instance}-A (IoU = 0.86), which is comparable to the original published results {[}\protect\hyperlink{ref-QmYuUQ5K}{11}{]}. For the 3D problem, the original backbone is not directly applicable, due to the aforementioned anisotropic issue and the images in the dataset do not contain enough Z-slices to run through all down sampling blocks in 3D. The anisotropic UNet {[}\protect\hyperlink{ref-jM3v1UjQ}{19}{]} is used here. The segmentation results obtained from the public dataset {[}\protect\hyperlink{ref-5sGcmDuy}{24}{]} contain nuclear instance segmentation of all cells. But, the cells touch the image borders are ignored from downstream analysis {[}\protect\hyperlink{ref-5sGcmDuy}{24}{]} and therefore not curated. In other words, the segmentation from this public dataset can only be used as high-quality nuclear instance segmentation ground truth after excluding the areas covered by cells touching the image borders {[}\protect\hyperlink{ref-5sGcmDuy}{24}{]}. Therefore, the exclusion masking function in \emph{MMV\_Im2Im} is very helpful in this example.

@@

We did a special comparison in this subsection to further illustrate the difference between semantic and instance segmentations. We took the 3D fibrillarin dataset from {[}\protect\hyperlink{ref-5sGcmDuy}{24}{]}. There are multiple channels in each 3D images, including DNA dye, membrane dye, and the structure channel (i.e., fibrillarin in this case). The original fibrillarin segmentation released with the dataset is a semantic segmentation (0=background, 1=fibrillarin). With the additional cell segmentation available in the dataset, we can know which groups of segmented fibrillarin belong to the same cell. Then, we can convert the original 3D fibrillarin segmentation segmentation ground truth into 3D instance segmentation ground truth (fibrillarin pixels belonging to the same cell are grouped as a unique instance). Sample images and results are shown in Figure \ref{fig:3dseg}. We can observe that the semantic segmentation model is able to achieve good accuracy in determining pixels from the fibrillarin signals. Meanwhile, the instance segmentation can group them properly so that fibrillarin masks from the same cell are successfully identified as unique instances, even without referring to the cell membrane channel or cell segmentation results. This is not a simple grouping step based on distance, since the fibrillarin signals from tightly touching nuclei may exist very close to each other.

\begin{figure}
\hypertarget{fig:3dseg}{%
\centering
\includegraphics[width=0.75\textwidth,height=\textheight]{images/semantic_seg3d.png}
\caption{Comparing 3D semantic segmentation and 3D instance segmentation results on confocal microscopy images of fibrillarin (showing a middle Z-slice of a 3D stack). From left to right: raw fibrillarin image, semantic segmentation, instance segmentation, instance ground truth, a reference images with all the channels (DNA dye in cyan, membrane dye in magenta, and fibrillarin in white) to show all the neighering cells.}\label{fig:3dseg}
}
\end{figure}

\hypertarget{unsupervised-semantic-segmentation-of-intracellular-structures-from-2d3d-confocal-microscopy-images}{%
\subsubsection{Unsupervised semantic segmentation of intracellular structures from 2D/3D confocal microscopy images}\label{unsupervised-semantic-segmentation-of-intracellular-structures-from-2d3d-confocal-microscopy-images}}

@@

\begin{longtable}[]{@{}llll@{}}
\caption{F1 scores of the unsupervised semantic segmentation predictions. \label{tbl:unsuper}}\label{tbl:unsuper}\tabularnewline
\toprule()
Dimensionality & Dataset & F1 Score & \# of Test Data \\
\midrule()
\endfirsthead
\toprule()
Dimensionality & Dataset & F1 Score & \# of Test Data \\
\midrule()
\endhead
2D & tight-junction & 0.888 ± 0.022 & 29 \\
3D & nucleus & 0.811 ± 0.150 & 15 \\
3D & golgi & 0.705 ± 0.022 & 6 \\
3D & mitochondria & 0.783 ± 0.005 & 2 \\
\bottomrule()
\end{longtable}

\begin{figure}
\hypertarget{fig:unsupervised}{%
\centering
\includegraphics[width=0.8\textwidth,height=\textheight]{images/unsupervised.png}
\caption{(A) Illustration of the unsupervised learning scheme and results in the 2D tight-junction segmentation problem. The yellow arrows indicate a region where the unsupervised learning method actually predicted better than the original segmentation results by classic image processing methods. (B) Example 3D segmentation results (only showing a middle z-slice) from models obtained by unsupervised learning.}\label{fig:unsupervised}
}
\end{figure}

\hypertarget{generating-synthetic-microscopy-images-from-binary-masks}{%
\subsubsection{Generating synthetic microscopy images from binary Masks}\label{generating-synthetic-microscopy-images-from-binary-masks}}

@@

For 2D demonstration, we extracted the middle Z-slice from NPM1 images as the training target, while using the NPM1 segmentation results as the input binary mask. With the paired ``mask + microscopy image'' data, we could train the model in a supervised fashion, or randomly shuffle the data to simulate the situation without paired data which can be trained in an unsupervised fashion. Example results can be found in Figure \ref{fig:synthetic}-A. In general, the supervised synthesization can generate more realistic images than the unsupervised model.

For 3D demonstration, we use 3D H2B images with two different types of input masks. First, we attempted to generate synthetic images from a coarse mask (i.e., only the overall shape of the nucleus, available as nuclear segmentation from the dataset) with both supervised training and unsupervised training. The unsupervised model in \emph{MMV\_Im2Im} uses the CycleGAN-based approaches. So, the unsupervised training is actually already done within the unsupervised segmentation experiments. In other words, the unsupervised model works in a bi-directional way, from real microscopy images to binary mask, and also from binary masks to simulated microscopy images. Here, we just do the inference in a different direction (from binary to simulated microscopy) using the model trained in unsupervised segmentation section. The results are shown in Figure \ref{fig:synthetic}-B (row 1). Similar to the results in 2D demonstration, the unsupervised synthesization can mostly ``paint'' the mask with homogeneous grayscale intensity, while the supervised model can simulate the textures to some extent. For a relatively large mask, it could be chanllenging for a model to fill in sufficient details to simulate real microscopy images (might be improved with diffusion-based models, see Discussions).

We made another attempt with 3D masks containing finer details beyond the overall shapes. So, we employed the H2B structure segmentation results from the dataset (capturing the detailed nuclear components marked by histon H2B) as the input for supervsied synthesization. The result is shown in Figure \ref{fig:synthetic}-B (row 2). Comparing to the synthesization with coarse masks, the images simulated from fine masks exhibit much more realistic appearance. As we can see, it is important to design the solutions with proper data.

\begin{figure}
\hypertarget{fig:synthetic}{%
\centering
\includegraphics[width=0.6\textwidth,height=\textheight]{images/synthetic.png}
\caption{Example results of (A) 2D synthetic fluorescent images of nucleoli (via NPM1) and (B) 3D synthetic fluorescent images of H2B (middle z-slices of a z-stack) with a coarse mask and a fine mask as the input.}\label{fig:synthetic}
}
\end{figure}

@@

\emph{MMV\_Im2Im} can also be used to computationally reduce image noise or restore the data from various sources of imaging artifacts, so as to increase the feasibility and efficiency in downstream analysis. In the current version of \emph{MMV\_Im2Im}, the restoration model can only be trained in a fully supervised manner. Therefore, aligned low quality and high quality images are required for supervision, even though such pair data can be partially simulated {[}\protect\hyperlink{ref-wcCVn8av}{6}{]}. Other methods, such as unsupervised learning based solutions {[}\protect\hyperlink{ref-4vnyY9J9}{41}{]}, will be made available within \emph{MMV\_Im2Im} in future versions.

>>>>

This manuscript
(\href{https://MMV-Lab.github.io/im2im-paper/v/95b4e4177071248e2ce1c3ffec17cc0c9df52191/}{permalink})
was automatically generated
from \href{https://github.com/MMV-Lab/im2im-paper/tree/95b4e4177071248e2ce1c3ffec17cc0c9df52191}{MMV-Lab/im2im-paper@95b4e41}
on September 26, 2023.

@@

\emph{Bottleneck: existing code not easy to understand or to extend or re-use}.

@@

Moreover, as ML scientists, have you ever overwhelmed by different training tricks for different methods or been curious about if certain state-of-the-art training methods can boost the performance of existing models? With the pytorch-lightning backend, \emph{MMV\_Im2Im} allows you to enjoy different state-of-the-art ML training engineering techniques without changing the code, e.g., stochastic weight averaging {[}\protect\hyperlink{ref-qzeQFRn9}{15}{]}, single precision training, automatic batch size determination, different optimizers, different learning rate schedulers, easy deployment on different devices, distributed training on multi-GPU (even multi-node), logging with common loggers such as Tensorboard, etc.. In short, with the pytorch-lightning based universal boilerplate, bioimaging researchers can really focus on research and develop novel methods for their biomedical applications, without worrying about the ML engineering works (which are usually lack in non-computer-science labs).

@@

The toolbox is designed for both people with or without extensive experience with ML and Python programming. It is not rare to find biomedical image analysis software that are very easy to use on a set of problems, but very hard to extend or adjust to other different but essentally related problems, or find some with great flexibility with tunable knobs at all levels, but unfortunately not easy for unexperienced users. To address this issue, we design the toolbox in a systematically modularized way with various levels of configurability. One can use the toolbox with a single command as simple as \texttt{run\_im2im\ -\/-config\ train\_semanticseg\_3d\ -\/-data.data\_path\ /path/to/data} or make customization on details directly from a human-readable configuration file, such as choosing batch normalization or instance normalization in certain layers of the model, or adding extra data augmentation steps, etc.. We provide an extensive list of more than 20 example configurations for various appplications and comprehensive documentation to address common questions for users as reference. For users preferring graphical interface, another napari plugin for MMV toolbox has been planned as the extension of \emph{MMV\_Im2Im} (see Discussion for details).

@@

Finally, we showed the comparison between three different types of models, an FCN-type model (i.e., fnet), a pix2pix-type model, and a cycleGAN-type model. For fair comparison, we used fnet as the same backbone for all three types of models. In theory, the pix2pix-type model can be trained in two different ways: from scratch or initializing the generator with a pre-trained fnet (trained as FCN). Examples of the comparison results were shown in the last two rows in Figure \ref{fig:labelfree}-B. Visually, it is evident that the additional adversarial components (i.e., the discriminator) could generate images with more realistic appearance than a typical FCN-type model alone, but again, we leave the appropriate quantitative evaluations to users' specfic biomedical studies.

@@

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2500}}@{}}
\caption{Evaluation of the final 3D label-free models for four different nuclear structures. \label{tbl:labelfree_table}}\label{tbl:labelfree_table}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Dataset
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Pearson Correlation
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Structural Similarity
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\# of Test Data
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Dataset
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Pearson Correlation
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Structural Similarity
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\# of Test Data
\end{minipage} \\
\midrule()
\endhead
FBL & 0.900 ± 0.013 & 0.858 ± 0.029 & 50 \\
HIST1H2BJ & 0.886 ± 0.023 & 0.758 ± 0.069 & 55 \\
LMNB1 & 0.884 ± 0.020 & 0.701 ± 0.061 & 50 \\
NPM1 & 0.937 ± 0.008 & 0.834 ± 0.025 & 55 \\
\bottomrule()
\end{longtable}

\begin{figure}
\hypertarget{fig:labelfree_comparison}{%
\centering
\includegraphics[width=0.76\textwidth,height=0.9\textheight]{images/labelfree_comparison_justin.png}
\caption{Examples of labelfree results. p/c/s refers percentile normalization, center normalization, and standard normalization, respectively (see main text for details).}\label{fig:labelfree_comparison}
}
\end{figure}

\begin{figure}
\hypertarget{fig:labelfree_comparison}{%
\centering
\includegraphics[width=0.76\textwidth,height=0.9\textheight]{images/labelfree_all_justin.png}
\caption{TODO.}\label{fig:labelfree_comparison}
}
\end{figure}

\hypertarget{d-semantic-segmentation-of-tissues-from-he-images-todo-update-figure-add-stds}{%
\subsubsection{2D semantic segmentation of tissues from H\&E images (TODO: Update figure, add stds)}\label{d-semantic-segmentation-of-tissues-from-he-images-todo-update-figure-add-stds}}

Segmentation is a common image processing task, and can be considered as a special type of image-to-image transformation, where the generated images are segmentation masks. DL-based methods have achieved huge success in semantic segmentation in biomedical images. In this example, we demonstrated \emph{MMV\_Im2Im} on a pathology application to segment glands from hematoxylin and eosin (H\&E) stained tissue images from the 2015 Gland Segmentation challenge {[}\protect\hyperlink{ref-45Sirz1X}{29},\protect\hyperlink{ref-XAffSYIR}{30}{]}. Stain normalization is an important pre-processing step in order to develop models robust to stain variation and tissue variations. \emph{MMV\_Im2Im} included a classic stain normalization method {[}\protect\hyperlink{ref-tQhnZyjK}{31}{]} as a pre-processing step. The effect of stain normalization can be observed in Figure \ref{fig:2d_gland}-A and B. We trained a simple attention UNet model {[}\protect\hyperlink{ref-OCow1hly}{26}{]}. Evaluated on the two different hold-out test sets, the model achieved F1-score, 0.904 ± 0.060 and 0.861 ± 0.117 on test set A and test set B, respectively. The performance was competitive compared to the methods reported in the challenge report {[}\protect\hyperlink{ref-XAffSYIR}{30}{]}, especially with much more consistent performance across the two different test sets. Example results can be found in Figure \ref{fig:2d_gland}-C.

\begin{figure}
\hypertarget{fig:2d_gland}{%
\centering
\includegraphics[width=0.75\textwidth,height=\textheight]{images/2d_semantic_seg_justin.png}
\caption{Example results of 2D semantic segmentation of gland in H\&E images.}\label{fig:2d_gland}
}
\end{figure}

@@

Instance segmentation is a type of segmentation problem that goes beyond semantic segmentation. The goal is to differentiate not only between different types of objects, but also different instances of the same type of objects. Currently, the \emph{MMV\_Im2Im} package supports \emph{EmbedSeg}-type models. The major benefit of EmbedSeg-type models is their agnosticism to the morphology and dimensionality of the object instances, compared to other models such as StarDist {[}\protect\hyperlink{ref-tIIG2f8K}{32},\protect\hyperlink{ref-14h90Vfg0}{33}{]}, SplineDist {[}\protect\hyperlink{ref-17Yrl6WGQ}{34}{]} and Cellpose {[}\protect\hyperlink{ref-TugPkOLy}{35}{]}. For example, different from the others, \emph{EmbedSeg}-type models are even able to generate instance segmentation where each instance contains multiple connected components. Additional frameworks such as Omnipose {[}\protect\hyperlink{ref-lXzmjM5n}{36}{]} will be supported in future versions. Another mainstream category of instance segmentation methods are detection-based models, such as Mask-RCNN {[}\protect\hyperlink{ref-xi8wnibR}{37}{]}. However, these models are better suited to the detection framework rather than image-to-image transformation (see Discussion section for details).

The \emph{EmbedSeg}-type models were re-implemented according to the original paper {[}\protect\hyperlink{ref-K2ugNcVa}{10},\protect\hyperlink{ref-QmYuUQ5K}{11}{]} following the generic boilerplate in \emph{MMV\_Im2Im}, with significant improvement. First of all, following the modular design of \emph{MMV\_Im2Im}, it is flexible to use different neural network models as the backbone. For 3D anisotropic microscopy image, the original backbone ERFNet {[}\protect\hyperlink{ref-XAkgs3Nh}{38}{]} doesn't take the anisotropic dimensions into account and therefore may not perform well or even not applicable. In this scenario, it is straighforward to employ another anisotropic neural network bone, such as the anisotropic U-Net in {[}\protect\hyperlink{ref-jM3v1UjQ}{19}{]} or the anisotropic version of Dynamic U-Net in MONAI. Second, we significantly improve training strategy. The original version requires pre-cropping patches centered around each instance and pre-calculated the center images and class images. This may generate a massive amount of additional data on the disk. More importantly, such pre-cropping makes data augmentation nearly impossible, except the simple ones like flipping (otherwise, the pre-calculated centers might be wrong), and also greatly undersamples around negative cases (e.g., background). For example, we have observed that for an Embedseg model training only with patches centered around instances, the model may suffer from degraded performance during inference when there are a large amount of background areas without any instances. Again, following the modular design of \emph{MMV\_Im2Im}, it is now possible to do on-the-fly data augmentation and patch sampling, even weighted patch sampling. Third, our improved \emph{Embedseg}-type models can accept an exclusion mask so that certain parts of the images can be ignored during training. This is especially useful for partially annotated ground truth. For large images, it could be extremely time-consuming to require every single instance to be annotated. The exclusion masks can address this bottleneck. Another extension comparing to the original implementation was that the \emph{MMV\_Im2Im} package made sliding windowing inference straightforward, and therefore permitted easy handling of images of any size during inference.

@@

For the 2D problem, we adopted the same network backbone as in the original \emph{EmbedSeg} paper. Example results on a small holdout set are shown in Figure \ref{fig:instance}-A (IoU = 0.879 ± 0.026), which is comparable to the original published results {[}\protect\hyperlink{ref-QmYuUQ5K}{11}{]}. For the 3D problem, the original backbone is not directly applicable, due to the aforementioned anisotropic issue and the images in the dataset do not contain enough Z-slices to run through all down sampling blocks in 3D. The anisotropic UNet {[}\protect\hyperlink{ref-jM3v1UjQ}{19}{]} is used here. The segmentation results obtained from the public dataset {[}\protect\hyperlink{ref-5sGcmDuy}{24}{]} contain nuclear instance segmentation of all cells. But, the cells touching the image borders are ignored from downstream analysis {[}\protect\hyperlink{ref-5sGcmDuy}{24}{]} and therefore not curated. In other words, the segmentation from this public dataset can only be used as high-quality nuclear instance segmentation ground truth after excluding the areas covered by cells touching the image borders {[}\protect\hyperlink{ref-5sGcmDuy}{24}{]}. Therefore, the exclusion masking function in \emph{MMV\_Im2Im} is very helpful in this example.

@@

We did a special comparison in this subsection to further illustrate the difference between semantic and instance segmentations. We took the 3D fibrillarin dataset from {[}\protect\hyperlink{ref-5sGcmDuy}{24}{]}. There are multiple channels in each 3D images, including DNA dye, membrane dye, and the structure channel (i.e., fibrillarin in this case). The original fibrillarin segmentation released with the dataset is a semantic segmentation (0=background, 1=fibrillarin). With the additional cell segmentation available in the dataset, we can know which groups of segmented fibrillarin belong to the same cell. Then, we can convert the original 3D fibrillarin semantic segmentation ground truth into 3D instance segmentation ground truth (fibrillarin pixels belonging to the same cell are grouped as a unique instance). Sample images and results are shown in Figure \ref{fig:3dseg}. We can observe that the semantic segmentation model is able to achieve good accuracy in determining pixels from the fibrillarin signals. Meanwhile, the instance segmentation can group them properly so that fibrillarin masks from the same cell are successfully identified as unique instances, even without referring to the cell membrane channel or cell segmentation results. This is not a simple grouping step based on distance, since the fibrillarin signals from tightly touching nuclei may exist very close to each other.

\begin{figure}
\hypertarget{fig:3dseg}{%
\centering
\includegraphics[width=0.75\textwidth,height=\textheight]{images/semantic_seg3d_justin.png}
\caption{Comparing 3D semantic segmentation and 3D instance segmentation results on confocal microscopy images of fibrillarin (showing a middle Z-slice of a 3D stack). From left to right: raw fibrillarin image, semantic segmentation, instance segmentation, instance ground truth, a reference images with all the channels (DNA dye in cyan, membrane dye in magenta, and fibrillarin in white) to show all the neighering cells.}\label{fig:3dseg}
}
\end{figure}

\hypertarget{unsupervised-semantic-segmentation-of-intracellular-structures-from-2d3d-confocal-microscopy-images-todo-adjust-text-2d-results-are-not-more-plausible-as-before}{%
\subsubsection{Unsupervised semantic segmentation of intracellular structures from 2D/3D confocal microscopy images (TODO: adjust text, 2d results are not more plausible as before)}\label{unsupervised-semantic-segmentation-of-intracellular-structures-from-2d3d-confocal-microscopy-images-todo-adjust-text-2d-results-are-not-more-plausible-as-before}}

@@

\begin{longtable}[]{@{}llll@{}}
\caption{F1 scores of the unsupervised semantic segmentation predictions. \label{tbl:unsuper}}\label{tbl:unsuper}\tabularnewline
\toprule()
Dimensionality & Dataset & F1 Score & \# of Test Data \\
\midrule()
\endfirsthead
\toprule()
Dimensionality & Dataset & F1 Score & \# of Test Data \\
\midrule()
\endhead
2D & tight-junction & 0.906 ± 0.011 & 18 \\
3D & nucleus & 0.836 ± 0.081 & 31 \\
3D & golgi & 0.689 ± 0.057 & 44 \\
3D & mitochondria & 0.804 ± 0.015 & 54 \\
\bottomrule()
\end{longtable}

\begin{figure}
\hypertarget{fig:unsupervised}{%
\centering
\includegraphics[width=0.8\textwidth,height=\textheight]{images/unsupervised_seg_justin.png}
\caption{(A) Illustration of the unsupervised learning scheme and results in the 2D tight-junction segmentation problem. The yellow arrows indicate a region where the unsupervised learning method actually predicted better than the original segmentation results by classic image processing methods. (B) Example 3D segmentation results (only showing a middle z-slice) from models obtained by unsupervised learning.}\label{fig:unsupervised}
}
\end{figure}

\hypertarget{generating-synthetic-microscopy-images-from-binary-masks}{%
\subsubsection{Generating synthetic microscopy images from binary masks}\label{generating-synthetic-microscopy-images-from-binary-masks}}

@@

For 2D demonstration, we extracted the middle Z-slice from NPM1 images as the training target, while using the NPM1 segmentation results as the input binary mask. With the paired ``mask + microscopy image'' data, we could train the model in a supervised fashion, or randomly shuffle the data to simulate the situation without paired data which can be trained in an unsupervised fashion using the CycleGAN framework implemented in \emph{MMV\_Im2Im}. Example results can be found in Figure \ref{fig:synthetic}-A. In general, the supervised synthesization can generate more realistic images than the unsupervised model.

For 3D demonstration, we use 3D H2B images with two different types of input masks. First, we attempted to generate synthetic images from a coarse mask (i.e., only the overall shape of the nucleus, available as nuclear segmentation from the dataset) with both supervised training and unsupervised training. The unsupervised model in \emph{MMV\_Im2Im} uses the CycleGAN-based approaches. So, the unsupervised training is actually already done within the unsupervised segmentation experiments. In other words, the unsupervised model works in a bi-directional way, from real microscopy images to binary masks, and also from binary masks to simulated microscopy images. Here, we just do the inference in a different direction (from binary to simulated microscopy) using the model trained in unsupervised segmentation section. The results are shown in Figure \ref{fig:synthetic}-B (row 1). Similar to the results in 2D demonstration, the unsupervised synthesization can mostly ``paint'' the mask with homogeneous grayscale intensity, while the supervised model can simulate the textures to some extent. For a relatively large mask, it could be chanllenging for a model to fill in sufficient details to simulate real microscopy images (might be improved with diffusion-based models, see Discussions).

We made another attempt with 3D masks containing finer details beyond the overall shapes. So, we employed the H2B structure segmentation results from the dataset (capturing the detailed nuclear components marked by histon H2B) as the input for supervsied synthesization. The result is shown in Figure \ref{fig:synthetic}-B (row 2). Compared to the synthesization with coarse masks, the images simulated from fine masks exhibit a much more realistic appearance. As we can see, it is important to design the solutions with proper data.

\begin{figure}
\hypertarget{fig:synthetic}{%
\centering
\includegraphics[width=0.6\textwidth,height=\textheight]{images/synthetic_justin.png}
\caption{Example results of (A) 2D synthetic fluorescent images of nucleoli (via NPM1) and (B) 3D synthetic fluorescent images of H2B (middle z-slices of a z-stack) with a coarse mask and a fine mask as the input.}\label{fig:synthetic}
}
\end{figure}

@@

\emph{MMV\_Im2Im} can also be used to computationally reduce image noise or restore the data from various sources of imaging artifacts, so as to increase the feasibility and efficiency in downstream analysis. In the current version of \emph{MMV\_Im2Im}, the restoration model can only be trained in a fully supervised manner. Therefore, aligned low-quality and high-quality images are required for supervision, even though such pair data can be partially simulated {[}\protect\hyperlink{ref-wcCVn8av}{6}{]}. Other methods, such as unsupervised learning-based solutions {[}\protect\hyperlink{ref-4vnyY9J9}{41}{]}, will be made available within \emph{MMV\_Im2Im} in future versions.
